The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-29 manuscript/manuscript.Rnw
Move more trees to supplemental

We examined the expression of three primates, \human, \bonobo, and
\smonkey. The genes whose expressions are different on the primate
lineage in comparison to all other analyzed species are listed in
\cref{tab:primate_lineage_specific}. \textit{FSTL1} has significantly
higher expression in the primate lineage
(\cref{fig:fstl1_evolution_tree},
$\mathrm{FDR}=\Sexpr{pri.all.lin.spec["FSTL1",`FDR`]}$), is implicated
in BMP signaling, and has previously been implicated in inflammation
and invasiveness
\cite{Mouillet.ea2015:expressionandpost-transcriptionalregulation}. Of
non-primates, \horse shows the highest expression
%
\Sexpr{gene.expression(species.name="equus caballus",gene="FSTL1",complete=TRUE)}, 
%
which is still lower than the lowest primate, \smonkey
%
(\Sexpr{gene.expression(species.name="ateles fusciceps",gene="FSTL1",complete=TRUE)}). 
%
No other genes were significantly differentially expressed on this
lineage by t-test.

\begin{figure}
<<fstl1.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="FSTL1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,min.fpkm=5)
@ 
\caption{\genetreeexplanation{follistatin-like 1 \textit{FSTL1}}}
\label{fig:fstl1_evolution_tree}
\end{figure}

@@

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \human are shown in
\cref{tab:top50_human_placenta} with all genes with \ac{FPKM} $\ge 50$
in Supplemental \cref{tab:all_homo_sapiens_placenta}. The gene with by far
the highest \ac{FPKM} value
%
(\Sexpr{gene.expression(species.name="homo sapiens",gene="CSH1")}) %
%
in \human is \textit{CSH1}, followed by \textit{CSH2}, and \textit{CSHL1}, all
three of which are members of a distinct expansion of the growth
hormone gene family within hominoids (\cref{fig:galectins_evolution_tree}). Another member of the growth
hormone gene family, \ac{GH2} (\ac{FPKM}=%
\Sexpr{gene.expression(species.name="homo sapiens",gene="GH2")}), %
is also expressed to a high degree in the \human placenta
(\cref{tab:top50_human_placenta}).
% 
Eleven \acp{PSG} are highly-expressed
(\ac{FPKM}$\ge 100$) in the term placenta of \human, with \textit{PSG1} showing the
greatest degree of expression %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG1")}) % 
and the pseudogene \textit{PSG10P} showing the least %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG10P")}, \cref{fig:psg_evolution_tree}) . %
These
genes are members of a hominoid-specific expansion of the \ac{CEA} family
and known to be expressed by the syncytiotrophoblast
\cite{Rawn.ea2008:evolutionregulationandfunction,Lei.ea1992:Cloningandexpressionof}. %
%
Genes involved in the production of estrogens (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP19A1",complete=TRUE)},
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD17B1",complete=TRUE)}, and
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)}%
% 
) and progesterone (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)},
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP11A1",complete=TRUE)}, and
\Sexpr{gene.expression(species.name="homo sapiens",gene="ADM",complete=TRUE)})
%
show exceptionally high expression in the human placenta, as well as %
%
\textit{CGA} (FPKM=\Sexpr{gene.expression(species.name="homo sapiens",gene="CGA")}, %
%
which generates the alpha subnit of multiple human chorionic
gonadotropin hormones (\cref{tab:top50_human_placenta}). The genes
producing the beta subunits of the chorionic gonadotropins
(\textit{CGB}, \textit{CGB1}, \textit{CGB2}, \textit{CGB5},
\textit{CGB7}, \textit{CGB8}) have reasonable expression
($\mathrm{FPKM} \ge \Sexpr{cgb.min.expression}$), and the sum of their
expression is on the order of the expression of expression of
\textit{CGA} ($\sum \mathit{CGBs}=\Sexpr{cgb.sum.expression}$ FPKM) .
%
%
Other highly expressed genes in the top 50 include members of the
hemoglobin family (e.g. \textit{HBA1}, \textit{HBB}, \textit{HBG2},
\textit{HBA2}), the maternally imprinted long noncoding RNA gene
\textit{H19}, genes coding for proteins required for growth and/or
cell proliferation (\textit{ADAM12}, \textit{KISS1}, \textit{IGF2},
\textit{GPC3}, \textit{CSF3R}), anticoagulation (\textit{TFPI2}), and
glucose transport (\textit{SLC2A1}) among others
(\cref{tab:top50_human_placenta}, Supplemental
\cref{tab:all_homo_sapiens_placenta}).

@@

\begin{table}
  \footnotesize
<<rodentia.lineage.specific,echo=FALSE,results="asis">>=
rod.all.lin.spec <- placenta.lineage.specific.table("rod.all")
setkey(rod.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed significantly in the Rodentia
  order as compared to all other analyzed species.}
\label{tab:rodentia_lineage_specific}
\end{table}

We examined the expression of three rodentia, \mouse, \carmeli, and
\galili{}. We lacked sufficient power to detect any significant
changes on the Rodentia lineage in comparison to all other analyzed
species (\cref{tab:rodentia_lineage_specific}).

@@

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_tile(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$ \ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

@@

In both ruminants and primates, the placenta is responsible for the
production of estrogen and progesterone. High relative expression of
\textit{CYP19A1} (\cref{fig:estrogen_evolution_tree}), the enzyme that
synthesizes estrogen from androstenedione is seen in \human, \bonobo,
\cow, and \horse, though not in \sheep, \smonkey. %
%
\textit{HSD3B1}, one of two enzymes that synthesizes progesterone from
pregnalone
(\cite{Simard.ea2005:Molecularbiologyof3beta-hydroxysteroid},
\cref{fig:progesterone_evolution_tree}) is also highly expressed in
\human, \bonobo, \cow, \horse, and \dog, but is not highly expressed
in \smonkey. The diversity of \textit{CYP19A1} and \textit{HSD3B1}
expression levels in in primates and ungulates seen here points to the
need for more complete sampling of placentas at different timepoints
and in different clades.

\begin{figure}
<<estrogen.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CYP19A1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{estrogen synthesis genes}}
\label{fig:estrogen_evolution_tree}
\end{figure}

\begin{figure}
<<progesterone.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="HSD3B1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSMODG00000005385","ENSLAFG00000027896")
                 )
@ 
\caption{\genetreeexplanation{progesterone synthesis genes}}
\label{fig:progesterone_evolution_tree}
\end{figure}

\begin{figure}
<<chorionic.somatomammotropin.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CSH1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{chorionic somatomammotropin}}
\label{fig:csh1_evolution_tree}
\end{figure}

@@

The elephant is strikingly the only mammal analyzed that shows high
expression of the primarily pituitary hormone prolactin (\textit{PRL})
in the placenta, although we see high expression of genes which
resulted from duplications of \textit{PRL} in the \mouse, \carmeli,
and \cow (\cref{fig:prolactin_evolution_tree}). All afrotherians yet
analyzed contain only one prolactin gene, and prolactin has undergone
adaptive evolution on the \elephant lineage
\cite{Wallis2009:ProlactininAfrotheriacharacterization}.
\citet{Yamamoto.ea2011:PlacentationinAfricanelephant} noted the
presence of a prolactin in the trophoblast of the elephant placenta
using immunohistological staining and human derived antibodies. A
previous analysis of \elephant placenta also detected the expression
of prolactin \cite{Hou.ea2012:Elephanttranscriptomeprovidesinsights}.
Here we confirm that the sole prolactin in \elephant is not only
expressed but highly expressed in the placenta, and likely functions
both in the maintenance of the corpus luteum and as a placental
lactogen \cite{Yamamoto.ea2011:PlacentationinAfricanelephant}.
\textit{PRL} is not expressed in \opossum, even though there is
evidence that it is expressed in the placenta of another marsupial,
\textit{Macropus eugenii} (tammar wallaby)
\cite{Menzies.ea2011:Placentalexpressionofpituitary}. This suggests
that placental prolactin expression in marsupials may be as diverse as
in eutherian mammals. One possible explanation of the diversity of
expression is the presence of endogenous retroviruses such as MER39 in
primates, MER77 in rodents, and L1-2\_LA in \elephant
\cite{Emera.ea2012:Convergentevolutionofendometrial} although this
does not explain the diversity of expression of different isoforms of
\textit{PRL} in \mouse or their tissue specificity in primates. An
alternative, that \textit{HOXA11} expression is responsible for
\textit{PRL} expression
\cite{Lynch.ea2008:Adaptivechangesintranscription} seems not to be the
case in placenta, as the rodent with the highest \textit{HOXA11}
expression, \galili
(\Sexpr{gene.expression("nannospalax_galili","Hoxa11",complete=TRUE)})
is the rodent with the lowest expression of \textit{PRL}
(\Sexpr{gene.expression("nannospalax_galili","Prl",complete=TRUE)},
\cref{fig:prolactin_evolution_tree}), and \elephant also does not
express \textit{HOXA11} highly either
(\Sexpr{gene.expression("loxodonta_africana","HOXA11",complete=TRUE)},
Supplemental \cref{fig:hoxa11_evolution_tree}).

\begin{figure}
<<prolactin.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{prolactin}}
\label{fig:prolactin_evolution_tree}
\end{figure}

@@

The opossum placenta shows expression of genes that either do not
exist in eutherians or are not found in the term eutherian placenta
(\cref{tab:top50_opossum_placenta}). The existence of embryonic forms
of hemoglobin in fetal tissues at term, as evidenced by the expression
of \textit{HBE} and \textit{HBZ} is a feature of marsupials that is
generally due to their very short gestation periods. Embryonic
hemoglobins in marsupials have a lower affinity for oxygen and allow
for a lower rate of oxygen uptake from maternal tissues
\cite{Holland.ea1997:Characterizationofembryonicglobin}. One novel
uncharacterized gene that shows high expression (\textit{LOC10001195}) is not
found in eutherians; Ensembl gene trees reveal that homologs of this
gene exist in other amniotes such as birds and turtles, but are not
present in eutherian mammals. The homolog of this gene in chicken is
found to expressed in reproductive tract
\cite{Froman.ea2006:expressedsequencetaganalysis}, but its function is
unknown. \textit{WFDC2} is a member of the whey acidic protein (WAP)
family and involved in innate immunity. It is expressed in the mammary
gland of the tammar wallaby during pregnancy and early lactation
\cite{Sharp.ea2007:Molecularevolutionofmonotreme}, and here in
relative abundance in the opossum placenta.

>>>>

We examined the expression of three primates, \human, \bonobo, and
\smonkey. The genes whose expressions are different on the primate
lineage in comparison to all other analyzed species are listed in
\cref{tab:primate_lineage_specific}. \textit{FSTL1} has significantly
higher expression in the primate lineage
(Supplemental \cref{fig:fstl1_evolution_tree},
$\mathrm{FDR}=\Sexpr{pri.all.lin.spec["FSTL1",`FDR`]}$), is implicated
in BMP signaling, and has previously been implicated in inflammation
and invasiveness
\cite{Mouillet.ea2015:expressionandpost-transcriptionalregulation}. Of
non-primates, \horse shows the highest expression
%
\Sexpr{gene.expression(species.name="equus caballus",gene="FSTL1",complete=TRUE)}, 
%
which is still lower than the lowest primate, \smonkey
%
(\Sexpr{gene.expression(species.name="ateles fusciceps",gene="FSTL1",complete=TRUE)}). 
%
No other genes were significantly differentially expressed on this
lineage by t-test.

@@

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \human are shown in
\cref{tab:top50_human_placenta} with all genes with \ac{FPKM} $\ge 50$
in Supplemental \cref{tab:all_homo_sapiens_placenta}. The gene with by far
the highest \ac{FPKM} value
%
(\Sexpr{gene.expression(species.name="homo sapiens",gene="CSH1")}) %
%
in \human is \textit{CSH1}, followed by \textit{CSH2}, and \textit{CSHL1}, all
three of which are members of a distinct expansion of the growth
hormone gene family within hominoids (\cref{fig:galectins_evolution_tree}). Another member of the growth
hormone gene family, \ac{GH2} (\ac{FPKM}=%
\Sexpr{gene.expression(species.name="homo sapiens",gene="GH2")}), %
is also expressed to a high degree in the \human placenta
(\cref{tab:top50_human_placenta}).
% 
Eleven \acp{PSG} are highly-expressed
(\ac{FPKM}$\ge 100$) in the term placenta of \human, with \textit{PSG1} showing the
greatest degree of expression %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG1")}) % 
and the pseudogene \textit{PSG10P} showing the least %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG10P")}, \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~A) . %
These
genes are members of a hominoid-specific expansion of the \ac{CEA} family
and known to be expressed by the syncytiotrophoblast
\cite{Rawn.ea2008:evolutionregulationandfunction,Lei.ea1992:Cloningandexpressionof}. %
%
Genes involved in the production of estrogens (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP19A1",complete=TRUE)},
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD17B1",complete=TRUE)}, and
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)}%
% 
) and progesterone (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)},
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP11A1",complete=TRUE)}, and
\Sexpr{gene.expression(species.name="homo sapiens",gene="ADM",complete=TRUE)})
%
show exceptionally high expression in the human placenta, as well as %
%
\textit{CGA} (FPKM=\Sexpr{gene.expression(species.name="homo sapiens",gene="CGA")}, %
%
which generates the alpha subnit of multiple human chorionic
gonadotropin hormones (\cref{tab:top50_human_placenta}). The genes
producing the beta subunits of the chorionic gonadotropins
(\textit{CGB}, \textit{CGB1}, \textit{CGB2}, \textit{CGB5},
\textit{CGB7}, \textit{CGB8}) have reasonable expression
($\mathrm{FPKM} \ge \Sexpr{cgb.min.expression}$), and the sum of their
expression is on the order of the expression of expression of
\textit{CGA} ($\sum \mathit{CGBs}=\Sexpr{cgb.sum.expression}$ FPKM) .
%
%
Other highly expressed genes in the top 50 include members of the
hemoglobin family (e.g. \textit{HBA1}, \textit{HBB}, \textit{HBG2},
\textit{HBA2}), the maternally imprinted long noncoding RNA gene
\textit{H19}, genes coding for proteins required for growth and/or
cell proliferation (\textit{ADAM12}, \textit{KISS1}, \textit{IGF2},
\textit{GPC3}, \textit{CSF3R}), anticoagulation (\textit{TFPI2}), and
glucose transport (\textit{SLC2A1}) among others
(\cref{tab:top50_human_placenta}, Supplemental
\cref{tab:all_homo_sapiens_placenta}).

@@

We examined the expression of three rodentia, \mouse, \carmeli, and
\galili{}. We lacked sufficient power to detect any significant
changes on the Rodentia lineage in comparison to all other analyzed
species (Supplemental \cref{tab:rodentia_lineage_specific}).

@@

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$ \ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

@@

In both ruminants and primates, the placenta is responsible for the
production of estrogen and progesterone. High relative expression of
\textit{CYP19A1} (Supplemental \cref{fig:estrogen_evolution_tree}),
the enzyme that synthesizes estrogen from androstenedione is seen in
\human, \bonobo, \cow, and \horse, though not in \sheep, \smonkey. %
%
\textit{HSD3B1}, one of two enzymes that synthesizes progesterone from
pregnalone
(\cite{Simard.ea2005:Molecularbiologyof3beta-hydroxysteroid},
Supplemental \cref{fig:progesterone_evolution_tree}) is also highly
expressed in \human, \bonobo, \cow, \horse, and \dog, but is not
highly expressed in \smonkey. The diversity of \textit{CYP19A1} and
\textit{HSD3B1} expression levels in in primates and ungulates seen
here points to the need for more complete sampling of placentas at
different timepoints and in different clades.

@@

The elephant is strikingly the only mammal analyzed that shows high
expression of the primarily pituitary hormone prolactin (\textit{PRL})
in the placenta, although we see high expression of genes which
resulted from duplications of \textit{PRL} in the \mouse, \carmeli,
and \cow (Supplemental \cref{fig:prolactin_evolution_tree}). All afrotherians yet
analyzed contain only one prolactin gene, and prolactin has undergone
adaptive evolution on the \elephant lineage
\cite{Wallis2009:ProlactininAfrotheriacharacterization}.
\citet{Yamamoto.ea2011:PlacentationinAfricanelephant} noted the
presence of a prolactin in the trophoblast of the elephant placenta
using immunohistological staining and human derived antibodies. A
previous analysis of \elephant placenta also detected the expression
of prolactin \cite{Hou.ea2012:Elephanttranscriptomeprovidesinsights}.
Here we confirm that the sole prolactin in \elephant is not only
expressed but highly expressed in the placenta, and likely functions
both in the maintenance of the corpus luteum and as a placental
lactogen \cite{Yamamoto.ea2011:PlacentationinAfricanelephant}.
\textit{PRL} is not expressed in \opossum, even though there is
evidence that it is expressed in the placenta of another marsupial,
\textit{Macropus eugenii} (tammar wallaby)
\cite{Menzies.ea2011:Placentalexpressionofpituitary}. This suggests
that placental prolactin expression in marsupials may be as diverse as
in eutherian mammals. One possible explanation of the diversity of
expression is the presence of endogenous retroviruses such as MER39 in
primates, MER77 in rodents, and L1-2\_LA in \elephant
\cite{Emera.ea2012:Convergentevolutionofendometrial} although this
does not explain the diversity of expression of different isoforms of
\textit{PRL} in \mouse or their tissue specificity in primates. An
alternative, that \textit{HOXA11} expression is responsible for
\textit{PRL} expression
\cite{Lynch.ea2008:Adaptivechangesintranscription} seems not to be the
case in placenta, as the rodent with the highest \textit{HOXA11}
expression, \galili
(\Sexpr{gene.expression("nannospalax_galili","Hoxa11",complete=TRUE)})
is the rodent with the lowest expression of \textit{PRL}
(\Sexpr{gene.expression("nannospalax_galili","Prl",complete=TRUE)},
\cref{fig:prolactin_evolution_tree}), and \elephant also does not
express \textit{HOXA11} highly either
(\Sexpr{gene.expression("loxodonta_africana","HOXA11",complete=TRUE)},
Supplemental \cref{fig:hoxa11_evolution_tree}).

@@

The \opossum placenta shows expression of genes that either do not
exist in eutherians or are not found in the term eutherian placenta
(\cref{tab:top50_opossum_placenta}). The existence of embryonic forms
of hemoglobin in fetal tissues at term, as evidenced by the expression
of \textit{HBE} and \textit{HBZ} is a feature of marsupials that is
generally due to their very short gestation periods. Embryonic
hemoglobins in marsupials have a lower affinity for oxygen and allow
for a lower rate of oxygen uptake from maternal tissues
\cite{Holland.ea1997:Characterizationofembryonicglobin}. One novel
uncharacterized gene that shows high expression (\textit{LOC10001195}) is not
found in eutherians; Ensembl gene trees reveal that homologs of this
gene exist in other amniotes such as birds and turtles, but are not
present in eutherian mammals. The homolog of this gene in chicken is
found to expressed in reproductive tract
\cite{Froman.ea2006:expressedsequencetaganalysis}, but its function is
unknown. \textit{WFDC2} is a member of the whey acidic protein (WAP)
family and involved in innate immunity. It is expressed in the mammary
gland of the tammar wallaby during pregnancy and early lactation
\cite{Sharp.ea2007:Molecularevolutionofmonotreme}, and present in
relative abundance in \opossum placenta.
